---
title: INTERCEPT interim analysis
author: "Joe Lewis"
date: today
date-format: "D MMMM YYYY"
format: beamer
execute:
  echo: false
  warning: false
  message: false
---

## Introduction

Aims of this analysis:

- Simple description of colonisation proportion by sample type stratified by
hospital/ward (where appropriate)
- Description of species stratified by sample type and resistance (CRO/ESBL)

Caveats:

- Based on chromagar only
- Not yet linked to REDcap metadata

```{r}

library(tidyverse)
library(here)
library(patchwork)

df <-
read_csv(here("data/processed/cleaned_inpatient_cultures.csv"))


```

```{r}
#| fig-width: 8
#| fig-height: 10


c <-
df |>
  select(hospital, sample_type, pid, day) |>
  unique() |>
  ggplot(aes(day)) +
  geom_bar() +
  facet_grid(hospital ~ sample_type) +
  labs(title = "Number of samples")
  
a <-
df |>
  select(hospital, sample_type, pid, day, plate, growth) |>
  group_by(hospital, sample_type, day, plate) |>
  summarise(
     n = sum(growth),
     N = length(growth)
) |>
rowwise() |>
mutate(prop = n/N,
       lci = binom.test(n,N)$conf.int[1],
       uci = binom.test(n,N)$conf.int[2]
) |>
filter(plate == "esbl") |>
  ggplot(
    aes(day, prop, ymin = lci, ymax = uci, group = plate, color = plate, fill =
plate)) + 
  geom_point(position = position_dodge(width = 1)) +
  geom_line( position = position_dodge(width = 1)) +
  #geom_errorbar(width = 0, position = position_dodge(width = 1)) + 
  geom_ribbon( position = position_dodge(width = 1), alpha = 0.3, color = NA) + 
  facet_grid(hospital ~ sample_type) +
  scale_fill_manual(values = "red") +
  scale_color_manual(values = "red") +
  labs(title = "Proportion of inpatients colonised: ESBL")

       
b <-
df |>
  select(hospital, sample_type, pid, day, plate, growth) |>
  group_by(hospital, sample_type, day, plate) |>
  summarise(
     n = sum(growth),
     N = length(growth)
) |>
rowwise() |>
mutate(prop = n/N,
       lci = binom.test(n,N)$conf.int[1],
       uci = binom.test(n,N)$conf.int[2]
) |>
filter(plate == "cro") |>
  ggplot(
    aes(day, prop, ymin = lci, ymax = uci, group = plate, color = plate, fill =
plate)) + 
  geom_point(position = position_dodge(width = 1)) +
  geom_line( position = position_dodge(width = 1)) +
  #geom_errorbar(width = 0, position = position_dodge(width = 1)) + 
  geom_ribbon( position = position_dodge(width = 1), alpha = 0.3, color = NA) + 
  facet_grid(hospital ~ sample_type) +
  scale_fill_manual(values = "blue") +
  scale_color_manual(values = "blue") +
  labs(title = "Proportion of inpatients colonised: CRO")

#c / a / b + plot_annotation(tag_levels = "A")



```

# Inpatients

## Inpatients: Number of samples


```{r}
#| fig-width: 5
#| fig-height: 4

c

```

## Inpatients: Proprotion colonised

```{r}
#| fig-width: 8
#| fig-height: 5

b + a

```

## Inpatients: species


```{r}
#| fig-width: 5
#| fig-height: 3

df |>
  separate_longer_delim(species, delim = ";") |>
  filter(!is.na(species), !grepl("SAMPLE|ABSENT", species)) |>
  mutate(species = gsub(",", "", species)) |>
  mutate(species = if_else(grepl("Coli|coli", species), "E. coli", species)) |>
  ggplot(aes(fct_infreq(species))) +
  geom_bar() +
  facet_wrap(~ plate, scales = "free_x") +
  coord_flip() +
  labs(x = "Species", title = "Number of isolates")

```

## Inpatients: CRO species vs time


```{r}
#| fig-width: 8
#| fig-height: 5

df |>
  filter(plate == "cro") |>
  group_by(hospital, sample_type, pid, day) |>
  separate_longer_delim(species, delim = ";") |>
  mutate(species = 
    case_when(
       grepl("Klebsiella", species) ~ "KEC",
       grepl("Coli|coli", species) ~ "E. coli",
       TRUE ~ species)) |>
  group_by(hospital, day, sample_type, species) |>
  count() |>
  filter(!is.na(species), !grepl("ABSENT|SAMPLE", species)) |>
  left_join(
    df |>
    select(hospital, sample_type, pid, day) |>
    unique() |> 
    group_by(hospital, sample_type, day) |> 
    count(name = "N"), 
    by = c("hospital", "sample_type", "day")
) |>
rowwise() |>
mutate(prop = n/N,
       lci = binom.test(n,N)$conf.int[1],
       uci = binom.test(n,N)$conf.int[2]
) |>
filter(species != "none") |>
  ggplot(
    aes(day, prop, ymin = lci, ymax = uci, color = species, fill = species)) + 
  geom_point(position = position_dodge(width = 1)) +
  geom_line( position = position_dodge(width = 1)) +
  #geom_errorbar(width = 0, position = position_dodge(width = 1)) + 
  geom_ribbon( position = position_dodge(width = 1), alpha = 0.3, color = NA) + 
  facet_grid(hospital ~ sample_type) 

```

# Environment


## Environmental swab results



```{r}
#| fig-width: 6
#| fig-height: 4


df <-
read_csv(here("data/processed/cleaned_env_cultures.csv"))

c <-
df |>
  select(hospital,  pid) |>
  unique() |>
  ggplot(aes(hospital)) +
  geom_bar() +
  labs(title = "Number of samples")
  
a <-
df |>
  select(hospital, pid, plate, growth) |>
  group_by(hospital, plate) |>
  summarise(
     n = sum(growth),
     N = length(growth)
) |>
rowwise() |>
mutate(prop = n/N,
       lci = binom.test(n,N)$conf.int[1],
       uci = binom.test(n,N)$conf.int[2]
) |>
  ggplot(
    aes(hospital, prop, ymin = lci, ymax = uci, group = plate, color = plate, fill =
plate)) + 
  geom_point(position = position_dodge(width = 0.6)) +
 # geom_line( position = position_dodge(width = 0.6)) +
  geom_errorbar(width = 0, position = position_dodge(width = 0.6)) + 
 # geom_ribbon( position = position_dodge(width = 1), alpha = 0.3, color = NA) + 
  scale_fill_manual(values = rev(c("red", "blue"))) + 
  scale_color_manual(values = rev((c("red", "blue")))) + 
  labs(title = "Proportion of swabs positive")

b <-
df |>
  separate_longer_delim(species, delim = ";") |>
  filter(!is.na(species), !grepl("SAMPLE|ABSENT|REJ", species)) |>
  mutate(species = gsub(",", "", species)) |>
  mutate(species = if_else(grepl("Coli|coli", species), "E. coli", species)) |>
  ggplot(aes(fct_infreq(species))) +
  geom_bar() +
  facet_wrap(~ plate, scales = "free_x") +
  coord_flip() +
  labs(x = "Species", title = "Number of isolates")

c + a

```

## Environmental species

```{r}
#| fig-width: 5
#| fig-height: 3

b
```

# Staff


## Staff swab results



```{r}
#| fig-width: 8
#| fig-height: 5


df <-
read_csv(here("data/processed/cleaned_staff_cultures.csv"))

c <-
df |>
  select(hospital,  pid, sample, day) |>
  unique() |>
  ggplot(aes(hospital, fill = sample)) +
  geom_bar(position = "dodge") +
  labs(title = "Number of samples") +
  facet_wrap( ~ day) +
  theme(legend.position = "bottom")
  
a <-
df |>
  select(hospital, pid, day, sample, plate, growth) |>
  group_by(hospital, plate, day, sample) |>
  summarise(
     n = sum(growth),
     N = length(growth)
) |>
rowwise() |>
mutate(prop = n/N,
       lci = binom.test(n,N)$conf.int[1],
       uci = binom.test(n,N)$conf.int[2]
) |>
mutate(day = gsub("swab", "", day)) |>
  ggplot(
    aes(day, prop, ymin = lci, ymax = uci, group = plate, color = plate, fill =
plate)) + 
  geom_point(position = position_dodge(width = 0.6)) +
  geom_line( position = position_dodge(width = 0.6)) +
  geom_errorbar(width = 0, position = position_dodge(width = 0.6)) + 
 # geom_ribbon( position = position_dodge(width = 1), alpha = 0.3, color = NA) + 
  scale_fill_manual(values = rev(c("red", "blue"))) + 
  scale_color_manual(values = rev((c("red", "blue")))) + 
  labs(title = "Proportion of swabs positive") +
  facet_grid(sample ~ hospital) +
  theme(legend.position = "bottom")

b <-
df |>
  separate_longer_delim(species, delim = ";") |>
  filter(!is.na(species), !grepl("SAMPLE|ABSENT|REJ", species)) |>
  mutate(species = gsub(",", "", species)) |>
  mutate(species = if_else(grepl("Coli|coli", species), "E. coli", species)) |>
  ggplot(aes(fct_infreq(species))) +
  geom_bar() +
  facet_grid(sample ~ plate, scales = "free_x") +
  coord_flip() +
  labs(x = "Species", title = "Number of isolates")

c + a

```

## Staff sample species

```{r}
#| fig-width: 8
#| fig-height: 5

b
```

# Community

## Community swab results


```{r}
#| fig-width: 6
#| fig-height: 4


df <-
read_csv(here("data/processed/cleaned_community_cultures.csv"))

c <-
df |>
  select(hospital,  pid) |>
  unique() |>
  ggplot(aes(hospital)) +
  geom_bar() +
  labs(title = "Number of samples")
  
a <-
df |>
  select(hospital, pid, plate, growth) |>
  group_by(hospital, plate) |>
  summarise(
     n = sum(growth),
     N = length(growth)
) |>
rowwise() |>
mutate(prop = n/N,
       lci = binom.test(n,N)$conf.int[1],
       uci = binom.test(n,N)$conf.int[2]
) |>
  ggplot(
    aes(hospital, prop, ymin = lci, ymax = uci, group = plate, color = plate, fill =
plate)) + 
  geom_point(position = position_dodge(width = 0.6)) +
 # geom_line( position = position_dodge(width = 0.6)) +
  geom_errorbar(width = 0, position = position_dodge(width = 0.6)) + 
 # geom_ribbon( position = position_dodge(width = 1), alpha = 0.3, color = NA) + 
  scale_fill_manual(values = rev(c("red", "blue"))) + 
  scale_color_manual(values = rev((c("red", "blue")))) + 
  labs(title = "Proportion of swabs positive")

b <-
df |>
  separate_longer_delim(species, delim = ";") |>
  filter(!is.na(species), !grepl("SAMPLE|ABSENT|REJ", species)) |>
  mutate(species = gsub(",", "", species)) |>
  mutate(species = if_else(grepl("Coli|coli", species), "E. coli", species)) |>
  ggplot(aes(fct_infreq(species))) +
  geom_bar() +
  facet_wrap(~ plate, scales = "free_x") +
  coord_flip() +
  labs(x = "Species", title = "Number of isolates")

c + a

```

## Community species


```{r}
#| fig-width: 5
#| fig-height: 3

b

```

## Conclusions

- High prevalence of ESBL colonsiation in community (86%) and in hospital.
- Low prevalance of CRO colonisation in community (6%).
- Increasing prevalence of CRO-colonisation in inpatients with time, mostly Acinetobacter.
- Widespread contamination of staff and environment with CRO resistant
organisms, most commonly Acinetobacter.

## Hypothesis - and intervention?

- Transmission of CRO-organmisms to patients occurs from staff and environment
- This can be interrupted with improved IPC 

## Further questions

- **What are drivers of differences in CRO prevalence between sites?**
  - Low CRO prevalence in Ashari ICU compared to Kariadi and Wongsonegoro; why?
    - Length of stay on ICU?
  - Higher CRO prevalence on ward at Kariadi compared to Wongsonegoro and
Ashari; why?
- **Role of antibiotics in driving increase in CROs**- needs addition of REDcap
data.
- Lots of **questions for genomics:**
  - Link environmental, staff, and patient isolates to infer transmission - does
this support hypothesis of transmission?
  - Are the same strains circulating in wrad environent causing invasive disease?
  - Are the same strains circulating in community and hospital?

# Next steps?
